## NCAA Football Attendance 

Let's load the tidyverse.

```{r}
library(tidyverse)
```

And read in the attendance data:

```{r}
attendance <- read_csv("https://dwillis.github.io/sports-data-files/attendance.csv")
```

Using the tidyverse's `pivot_longer` function, turn the currently wide data into longer data and save it as a new dataframe called `attendance_longer`:

```{r}
attendance_longer <- attendance |> pivot_longer(cols = starts_with("20"), names_to = "Year", values_to = "Attendance")

attendance_longer |> filter(Attendance >= 500000)
```

Let's try to figure out whether football attendance has returned to "normal" by 2022, defining "normal" as 2019 figures. For that we'll need to isolate those two years using `filter()` and get rid of any NAs:

```{r}
attendance_two_years <- attendance_longer |> filter(Year == "2019" | Year == "2022") |> na.omit()
```

Now, in order to calculate a percentage change between those two years, we'll need to go back to wider data and again ditch teams where we don't have both years:

```{r}
attendance_wide <- attendance_two_years |> pivot_wider(names_from = Year, values_from = Attendance) |> na.omit()
```

Finally, we'll add a percentage change column to that dataframe (remember, it's (New-Old)/Old) and arrange it to see the results.

```{r}
attendance_wide_pct <- attendance_wide |> mutate(
  pct_change = ((`2022` - `2019`)/`2019`)*100
) |> arrange(desc(pct_change))

attendance_wide_conference <- attendance_wide |>
  group_by(Conference) |> 
  summarise(
    total_2019 = sum(`2019`),
    total_2022 = sum(`2022`)
  ) |> mutate(
  pct_change = ((total_2022 - total_2019)/total_2019)*100) |> 
  arrange(desc(pct_change))
```

Some of these clearly are influenced by performance on the field. Below, describe the most interesting results (positive or negative) and what you think might be going on with them. Feel free to do some googling and describe what other data would be helpful to clarify what the best stories here might be. What surprises you about these results?

**ANSWER**
From 2019 to 2022, the institution with the largest positive change in attendance from 2019 to 2022 is UTSA, with an increase of 69.497%, while the institution with the largest negative change in attendance from the same period is Hawaii, with a decrease of -69.459%. Hawaii is not part of the mainland United States, likely making it harder for opposing teams to visit for scheduled games, especially during the COVID-19 pandemic. I do suspect if the greater change in attendance for UTSA was either due to increased admissions or how the university rolled out their COVID-19 policies compared to other universities at the time (or maybe a combination of both).

If we consider adding all of the attendances by conference and comparing any changes (attendance_wide_conference), Big 12 had the greatest increase from that period at 5.734% whereas MAC had the greatest decrease at -9.268%. However, I don't think these changes mean anything significant, as I would consider a change of 10% or greater to indicate something I would have to look deeper into.
