```{r}
library(tidyverse)
```

```{r}
volleyball_matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2023.csv")
```

```{r}
glimpse(volleyball_matches_2023)
```

```{r}
perset <- volleyball_matches_2023 |> 
  select(date, team, opponent, total_attacks, assists, block_solos, block_assists, defensive_total_attacks, defensive_assists, defensive_block_solos, defensive_block_assists)
```

```{r}
teamtotals <- perset |>
  group_by(team) |> 
   summarise(
    AttacksAvg = mean(total_attacks), 
    AssistsAvg = mean(assists), 
    BlocksAvg = mean(block_solos + block_assists),
    DefAttacksAvg = mean(defensive_total_attacks), 
    DefAssistsAvg = mean(defensive_assists), 
    DefBlocksAvg = mean(defensive_block_solos + defensive_block_assists)
    ) 

teamzscore <- teamtotals |> 
  mutate(
    AttacksZscore = as.numeric(scale(AttacksAvg, center = TRUE, scale = TRUE)),
    AssistsZscore = as.numeric(scale(AssistsAvg, center = TRUE, scale = TRUE)),
    BlocksZscore = as.numeric(scale(BlocksAvg, center = TRUE, scale = TRUE)),
    DefAttacksZscore = as.numeric(scale(DefAttacksAvg, center = TRUE, scale = TRUE)) * -1,
    DefAssistsZScore = as.numeric(scale(DefAssistsAvg, center = TRUE, scale = TRUE)) * -1,
    DefBlocksZScore = as.numeric(scale(DefBlocksAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = AttacksZscore + AssistsZscore + BlocksZscore + DefAttacksZscore + DefAssistsZScore + DefBlocksZScore
  )  
```

```{r}
head(teamzscore)
```
```{r}
teamzscore |> 
  arrange(desc(TotalZscore)) |>
  select(team, TotalZscore)
```

The Oregon Ducks are currently the top performing college volleyball team based on their attacks, assists and blocks achieved during game play.

The results were determined from a database of college volleyball matches for the Fall 2023 season. Compared to other teams, the Ducks have an average of 116.154 attacks during matches, putting them among the top 80% of teams based on their average number of attacks. The Ducks also have an average of 47.077 in during matches, putting them among the top 30 teams. Finally, they have an average of exactly 19 blocks (combine solo and assists), putting them among the top 10 teams based off this measurement alone.

As of September 24th (after 13 games), the Ducks have a win-loss ratio of 12:1. Other teams that performed as well as the Ducks are the Stanford Cardinal, the Arizona St. Sun Devils, and the Towson Tigers. The team currently performing the worse are the Le Moyne Dolphins, as all of their (currently) 12 games are losses. Other teams that performed as worse as the Dolphins are the Alcorn Lady Braves, the Manhattan Jaspers, and the Southern Utah Thunderbirds.

Perhaps one reason why the Oregon Ducks are performing well in volleyball is because they are based in Eugene, which is known by many as "Track Town, USA" because of the university's successful football and track & field programs. In addition, the renowned athletic footwear and apparel company Nike being founded by alumni Phil Knight and Bill Bowerman. According to the University of Oregon's website, their annual operating budget is more than \$1.1 billion, with a considerable portion of it covering auxiliary operations, such as athletics. The university also mentions that their athletics department is self-sufficient, meaning it receives no tuition or state operating funds, as the department paid around \$15 million for both administrative overhead and student-athlete scholarships.

https://www.uoregon.edu/accountability/transparency/faq

```{r}
rosters <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-volleyball/main/vb_rosters_2023_24.csv")
```
```{r}
rosters_with_towns <- rosters |>
  mutate(HometownCopy = Hometown) |>
  separate(Hometown, c('Town', 'State'), sep=',') |>
  separate(Height, c('Feet', 'Inches'), sep=' ')
```
Dumbbell and lollipop charts

```{r}
logs <- read_csv("https://dwillis.github.io/sports-data-files/ncaa_womens_volleyball_matchstats_2022.csv")
```
```{r}
library(tidyverse)
library(ggalt)
```
```{r}
big10 <- c("Nebraska Cornhuskers", "Iowa Hawkeyes", "Minnesota Golden Gophers", "Illinois Fighting Illini", "Northwestern Wildcats", "Wisconsin Badgers", "Indiana Hoosiers", "Purdue Boilermakers", "Ohio St. Buckeyes", "Michigan Wolverines", "Michigan St. Spartans", "Penn St. Nittany Lions", "Rutgers Scarlet Knights", "Maryland Terrapins, Terps")

errors <- logs |>
  filter(team %in% big10) |> 
  group_by(team) |> 
  summarise(
    total_errors = sum(errors), 
    opp_errors = sum(defensive_errors))
```

```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=team, x=total_errors, xend=opp_errors)
  )
```
```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=team, x=total_errors, xend=opp_errors),
    size=1,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red")
```
```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=total_errors, xend=opp_errors),
    linewidth = 1,
    size=1,
    color = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal()
```
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=team, x=opp_errors), 
    horizontal = TRUE,
    )
```
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=opp_errors), 
    horizontal = TRUE
    ) + theme_minimal() + 
  labs(title = "Maryland, Rutgers force more errors", y="team")
```
```{r}
md <- errors |> filter(team == "Maryland Terrapins, Terps")

ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=opp_errors), 
    horizontal = TRUE
    ) + 
  geom_lollipop(
    data=md,
    aes(y=team, x=opp_errors),
    horizontal = TRUE,
    color = "red"
  ) + 
  theme_minimal() + 
  labs(title = "Maryland forced the most errors among Big Ten teams", y="team")
```

