---
title: "First Presentation - Data Backgrounding"
author: "Mitchell Hang"
date: '2023-10-07'
format: 
  html:
    code-fold: true
---

The question that I have proposed for this project is...

"Who is considered to be the top tennis player in 2022, based on their appearances and rankings in any of the four Grand Slams (Australian Open, French Open, Wimbledon, U.S. Open) as well as other competitive ATP/WTA tournaments, based on cumulative points/sets/matches won?"

I decided to go for this question because many people have considered who the greatest tennis player currently is, but few have taken the time to calculate such findings based on overall performance metrics. I feel that it's better that we answer this question based on objectively compiling and analyzing competition results rather than subjective opinions and thinking, because people will get an idea of how one player performs compared to another.

In order to tackle this question, I would have to search for tennis databases on the web. The data in question can be found on multiple sites, particularly those hosted by the Association of Tennis Professionals (ATP) and the Women's Tennis Association (WTA), as well as other secondary sites such as ultimatetennisstatistics.com, tennisexplorer.com, and tennisabstract.com. For instance, the latter site is very helpful because it gives me the opportunity to filter the database based on the kind of court played (grass, clay, hard), the specific match/tournament, and their age. However, the database seems to include only the top 50 players per category, so I would consider finding a database that compiles as many registered players that competed in tournaments throughout 2022. I would also sub-sort/filter the data by country, age (group) and tournament(s) they competed in in case I would have to find any sub-totals. 

For simplification purposes, I am focusing solely on tennis players who competed in 2022. The fourth Grand Slam (U.S. Open) is already over for this year (2023), but I would rather stick to 2022 because there are some other competitions in-progress that other players are competing in, and we currently don't know of the results yet. It seems someone on GitHub has already compiled a database of ATP matches for the past few decades, so I may consider basing my current work off of that (with the appropriate credit).

```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(ggalt)
```

Database found here (2023 data is up to September 11, 2023 as of writing this): 
https://github.com/JeffSackmann/tennis_atp/blob/master/atp_matches_2022.csv
https://github.com/JeffSackmann/tennis_atp/blob/master/atp_matches_2023.csv
https://github.com/JeffSackmann/tennis_wta/blob/master/wta_matches_2022.csv
https://github.com/JeffSackmann/tennis_wta/blob/master/wta_matches_2023.csv

```{r}
atp_matches_2022 <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_matches_2022.csv")

atp_matches_2023 <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_matches_2023.csv")

wta_matches_2022 <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_wta/master/wta_matches_2022.csv")

wta_matches_2023 <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_wta/master/wta_matches_2023.csv")
```

```{r}
atp_matches_2022_new <- atp_matches_2022 |>
  mutate(tourney_date = ymd(tourney_date))

atp_matches_2023_new <- atp_matches_2023 |>
  mutate(tourney_date = ymd(tourney_date))

wta_matches_2022_new <- wta_matches_2022 |>
  mutate(tourney_date = ymd(tourney_date))

wta_matches_2023_new <- wta_matches_2023 |>
  mutate(tourney_date = ymd(tourney_date))

grandslams <- c("Wimbledon", "Us Open", "Roland Garros", "Australian Open")
```

Top 20 Male Tennis Players across the Four Grand Slams in 2022 (based on total matches won)
```{r}
atp_matches_2022_new_gs <- atp_matches_2022_new |>
  filter(tourney_name %in% grandslams) |>
  group_by(winner_name) |>
  summarise(matches_won = n()) |>
  arrange(desc(matches_won)) |>
  head(20)

atp_matches_2022_new_gs
```
```{r}
ggplot() + 
  geom_bar(
    data=atp_matches_2022_new_gs, 
    aes(x=reorder(winner_name, matches_won),
        weight=matches_won)) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Rafael Nadal won the most men's Grand Slam matches for 2022", 
    x="Player", 
    y="Matches won") + 
  theme_minimal() + 
  coord_flip()
```

Top 20 Male Tennis Players across the Four Grand Slams in 2023 (based on total matches won)
```{r}
atp_matches_2023_new_gs <- atp_matches_2023_new |>
  filter(tourney_name %in% grandslams) |>
  group_by(winner_name) |>
  summarise(matches_won = n()) |>
  arrange(desc(matches_won)) |>
  head(20)

atp_matches_2023_new_gs
```
```{r}
ggplot() + 
  geom_bar(
    data=atp_matches_2023_new_gs, 
    aes(x=reorder(winner_name, matches_won),
        weight=matches_won)) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Novak Djokovic won the most men's Grand Slam matches for 2023", 
    x="Player", 
    y="Matches won") + 
  theme_minimal() + 
  coord_flip()
```

Top 20 Female Tennis Players across the Four Grand Slams in 2022 (based on total matches won)
```{r}
wta_matches_2022_new_gs <- wta_matches_2022_new |>
  filter(tourney_name %in% grandslams) |>
  group_by(winner_name) |>
  summarise(matches_won = n()) |>
  arrange(desc(matches_won)) |>
  head(20)

wta_matches_2022_new_gs
```
```{r}
ggplot() + 
  geom_bar(
    data=wta_matches_2022_new_gs, 
    aes(x=reorder(winner_name, matches_won),
        weight=matches_won)) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Iga Swiatek won the most women's Grand Slam matches for 2022", 
    x="Player", 
    y="Matches won") + 
  theme_minimal() + 
  coord_flip()
```

Top 20 Female Tennis Players across the Four Grand Slams in 2023 (based on total matches won)
```{r}
wta_matches_2023_new_gs <- wta_matches_2023_new |>
  filter(tourney_name %in% grandslams) |>
  group_by(winner_name) |>
  summarise(matches_won = n()) |>
  arrange(desc(matches_won)) |>
  head(20)

wta_matches_2023_new_gs
```

```{r}
ggplot() + 
  geom_bar(
    data=wta_matches_2023_new_gs, 
    aes(x=reorder(winner_name, matches_won),
        weight=matches_won)) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Aryna Sabalenka won the most women's Grand Slam matches for 2023", 
    x="Player", 
    y="Matches won") + 
  theme_minimal() + 
  coord_flip()
```

So far from my initial explorations in the datasets (I added two more for the WTA from 2022 and 2023), I was able to construct analyses (both tibbles and bar graphs) for the top 20 male and female players (because 10 was too small) across the four Grand Slams in both 2022 and 2023, based on the amount of matches they won. Rafael Nadal and Iga Swiatek won the most Grand Slam matches in 2022, wheras Novak Djokovic and Aryna Sabalenka won the most Grand Slam matches in 2023. Interestingly enough, those four players have won more than 20 Grand Slam matches whereas many other players have won fewer than 20. It'll be interesting to compare all of the players in 2022 and 2023 by selecting all of the matches outside of the Grand Slams to see if anyone remains on top or if we see a sudden rise or fall in the rankings.

I noticed that both Novak Djokovic and Aryna Sabalenka made surprising jumps in the ranking from 2022 to 2023, and I was surprised to see Rafael Nadal in the Top 20 male ranking for 2023, while checking the dataset reveals that he only won one Grand Slam match (and further research revealed that he had suffered an injury at the Australian Open and subsequently did not appear in the subsequent Grand Slams), so I was really surprised to see Rafael Nadal have what is considerably the largest drop in number of Grand Slam matches won from 2022 and 2023. 

Upon closer inspection of these datasets, they list a timeline of the matches which have occurred and other statistics for both players such as their aces, total serve points, and break points saved/faced. Since I currently don't know how to efficiently tally up without any errors, I feel that it's much easier on matches won rather than sets or points. Alternatively, I can make a line chart out of the "winner rank" column showing a player's progression over time, and I might take the top 20 (or maybe 10 since 20 might be too many) players and give them lines of different colors while I put the rest of the players in grey (sorta like how I did with Lesson #15).

-----

Here I am creating a cluster based on the top 20 male and female winners of the Grand Slams for 2022 and 2023.

```{r}
male_winners_2022 <- c("Rafael Nadal", "Carlos Alcaraz", "Jannik Sinner", "Casper Ruud", "Daniil Medvedev", "Marin Cilic", "Nick Kyrgios", "Novak Djokovic", "Andrey Rublev", "Cameron Norrie", "Frances Tiafoe", "Karen Khachanov", "Stefanos Tsitsipas", "Cristian Garin", "Matteo Berrettini", "Alex De Minaur", "Alexander Zverev", "Botic Van De Zandschulp", "Felix Auger Aliassime", "Miomir Kecmanovic")

male_winners_2023 <- c("Novak Djokovic","Carlos Alcaraz","Andrey Rublev","Stefanos Tsitsipas","Daniil Medvedev", "Alexander Zverev", "Jannik Sinner", "Holger Rune", "Tommy Paul", "Ben Shelton", "Frances Tiafoe", "Grigor Dimitrov", "Casper Ruud", "Hubert Hurkacz", "Karen Khachanov", "Alex De Minaur", "Jiri Lehecka", "Nicolas Jarry", "Taylor Fritz", "Alejandro Davidovich Fokina")
  
female_winners_2022 <- c("Iga Swiatek", "Jessica Pegula", "Cori Gauff", "Ons Jabeur", "Alize Cornet", "Amanda Anisimova", "Aryna Sabalenka", "Danielle Collins", "Elena Rybakina", "Madison Keys", "Ajla Tomljanovic", "Caroline Garcia", "Elise Mertens", "Paula Badosa", "Simona Halep", "Veronika Kudermetova", "Qinwen Zheng", "Victoria Azarenka", "Alison Riske Amritraj", "Ashleigh Barty")

female_winners_2023 <- c("Aryna Sabalenka", "Iga Swiatek", "Cori Gauff", "Elena Rybakina", "Marketa Vondrousova", "Ons Jabeur", "Jessica Pegula", "Karolina Muchova", "Madison Keys", "Elina Svitolina", "Jelena Ostapenko", "Beatriz Haddad Maia", "Belinda Bencic", "Ekaterina Alexandrova", "Victoria Azarenka", "Daria Kasatkina", "Elise Mertens", "Magda Linette", "Bernarda Pera", "Donna Vekic")
```

Here I am creating a dumbbell chart showing the overall match wins and losses for the Top 20 male Grand Slam winners in 2022. Rafael Nadal, the top Grand Slam winner in 2022, is surprisingly not ranked at the top of the chart but somewhere towards the middle, as the number of wins and losses (as well as the difference between the two) comes close to the 2023 Grand Slam winner, Novak Djokovic. I was quite surprised to see Cristain Garin as being the only player who has had more match losses than wins in 2022, and he also had the least number of match wins that year. 

```{r}
atp_matches_2022_new_top20_win <- atp_matches_2022_new |>
  filter(winner_name %in% male_winners_2022) |>
  group_by(winner_name) |>
  rename(name = winner_name) |>
  summarise(matches_won = n())

atp_matches_2022_new_top20_lose <- atp_matches_2022_new |>
  filter(loser_name %in% male_winners_2022) |>
  group_by(loser_name) |>
  rename(name = loser_name) |>
  summarise(matches_lost = n())

atp_matches_2022_new_top20_winlose <- atp_matches_2022_new_top20_win |> left_join(atp_matches_2022_new_top20_lose)

ggplot() + 
  geom_dumbbell(
    data=atp_matches_2022_new_top20_winlose, 
    aes(y=reorder(name, matches_won), x=matches_won, xend=matches_lost),
    linewidth = 1,
    size=1,
    color = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal() + 
  labs(title="Alcaraz has largest win-loss gap in 2022", subtitle="Stefanos Tsitsipas had the highest overall match wins in 2022, but Carlos Alcaraz had the \nhighest largest difference between match wins and losses during the same year \n(green = matches won, blue = matches lost).", caption="Source: ATP | By Mitchell Hang")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```
Here I am creating a dumbbell chart showing the overall match wins and losses for the Top 20 male Grand Slam winners in 2023. Carlos Alcaraz, who was third in the 2022 chart, is now at the top of the 2023 chart, implying that he is a top performer who has improved over the past year in his match wins. Novak Djokovic was the top Grand Slam winner in 2023, and he is also climbing up the chart, as he was near the middle of the ranking from last year. Like in 2022, there was only one player who had more losses than wins, and this time it's Ben Shelton.

```{r}
atp_matches_2023_new_top20_win <- atp_matches_2023_new |>
  filter(winner_name %in% male_winners_2023) |>
  group_by(winner_name) |>
  rename(name = winner_name) |>
  summarise(matches_won = n())

atp_matches_2023_new_top20_lose <- atp_matches_2023_new |>
  filter(loser_name %in% male_winners_2023) |>
  group_by(loser_name) |>
  rename(name = loser_name) |>
  summarise(matches_lost = n())

atp_matches_2023_new_top20_winlose <- atp_matches_2023_new_top20_win |> left_join(atp_matches_2023_new_top20_lose)

ggplot() + 
  geom_dumbbell(
    data=atp_matches_2023_new_top20_winlose, 
    aes(y=reorder(name, matches_won), x=matches_won, xend=matches_lost),
    linewidth = 1,
    size=1,
    color = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal() + 
  labs(title="Alcaraz's win-loss gap grows larger in 2023", subtitle="Carlos Alcaraz now takes the title of having the highest number of match wins in 2023 \nas his win-loss gap grows larger, while other players such as 2023 Grand Slam top winner \nNovak Djokovic is climbing up the rankings (green = matches won, red = matches lost).", caption="Source: ATP | By Mitchell Hang")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```
Here I am creating a dumbbell chart showing the overall match wins and losses for the Top 20 female Grand Slam winners in 2022. It's clear to see that Iga Swiatek has dominated much of the competition in the WTA during that year, as she not only has the highest number of match wins but also the largest difference between the number of match wins and losses. Despite it looking like that Swiatek has the fewest match losses out of the top 20, it's actually Ashleigh Barty which has the fewest losses at 0, meaning that she won every single one of her matches she played during that year.

```{r}
wta_matches_2022_new_top20_win <- wta_matches_2022_new |>
  filter(winner_name %in% female_winners_2022) |>
  group_by(winner_name) |>
  rename(name = winner_name) |>
  summarise(matches_won = n())

wta_matches_2022_new_top20_lose <- wta_matches_2022_new |>
  filter(loser_name %in% female_winners_2022) |>
  group_by(loser_name) |>
  rename(name = loser_name) |>
  summarise(matches_lost = n())

wta_matches_2022_new_top20_winlose <- wta_matches_2022_new_top20_win |> left_join(wta_matches_2022_new_top20_lose) 

wta_matches_2022_new_top20_winlose <- wta_matches_2022_new_top20_winlose |> replace(is.na(wta_matches_2022_new_top20_winlose), 0)

ggplot() + 
  geom_dumbbell(
    data=wta_matches_2022_new_top20_winlose, 
    aes(y=reorder(name, matches_won), x=matches_won, xend=matches_lost),
    linewidth = 1,
    size=1,
    color = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal() + 
  labs(title="Swiatek has largest win-loss gap in 2022", subtitle="Iga Swiatek has not only the highest number of match wins in 2022, but also the largest \ndifference between match wins and losses during that same year (green = matches won, \nred = matches lost).", caption="Source: WTA | By Mitchell Hang")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```
Here I am creating a dumbbell chart showing the overall match wins and losses for the Top 20 female Grand Slam winners in 2023. Like in 2022, Iga Swiatek's gap between match wins and losses is relatively large, even though the numbers for both are slightly smaller for this year. The 2023 top Grand Slam winner, Aryna Sabalenka, is right below Swiatek on the chart, as she has a smaller number of match wins but the same number of match losses as Swiatek. This time, there was one player who had more losses than wins, and that is Bernada Pera, but even she doesn't have the lowest number of match wins, as that belongs to Elina Svitolina, who also has the lowest number of match losses.

```{r}
wta_matches_2023_new_top20_win <- wta_matches_2023_new |>
  filter(winner_name %in% female_winners_2023) |>
  group_by(winner_name) |>
  rename(name = winner_name) |>
  summarise(matches_won = n())

wta_matches_2023_new_top20_lose <- wta_matches_2023_new |>
  filter(loser_name %in% female_winners_2023) |>
  group_by(loser_name) |>
  rename(name = loser_name) |>
  summarise(matches_lost = n())

wta_matches_2023_new_top20_winlose <- wta_matches_2023_new_top20_win |> left_join(wta_matches_2023_new_top20_lose)

ggplot() + 
  geom_dumbbell(
    data=wta_matches_2023_new_top20_winlose, 
    aes(y=reorder(name, matches_won), x=matches_won, xend=matches_lost),
    linewidth = 1,
    size=1,
    color = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal() + 
  labs(title="Swiatek's win-loss gap shrinks in 2023", subtitle="After Iga Swiatek had the largest difference between match wins and losses in 2022, she \ncontinuesto hold the highest number of match wins in 2023, but the win-loss gap has gotten \nslightly smaller (green = matches won, red = matches lost).", caption="Source: WTA | By Mitchell Hang")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```

##### Z-scores

```{r}
atp_matches_2022_selected <- atp_matches_2022 |>
  select(winner_name, w_ace, w_df, w_svpt, l_ace, l_df, l_svpt)

atp_matches_2023_selected <- atp_matches_2023 |>
  select(winner_name, w_ace, w_df, w_svpt, l_ace, l_df, l_svpt)

wta_matches_2022_selected <- wta_matches_2022 |>
  select(winner_name, w_ace, w_df, w_svpt, l_ace, l_df, l_svpt)

wta_matches_2023_selected <- wta_matches_2023 |>
  select(winner_name, w_ace, w_df, w_svpt, l_ace, l_df, l_svpt)
```

```{r}
atp_matches_2022_totals <- atp_matches_2022_selected |> 
  group_by(winner_name) |> 
  na.omit() |>
  summarise(
    winnerMeanAce = mean(w_ace), 
    winnerMeanDf = mean(w_df), 
    winnerMeanSvpt = mean(w_svpt),
    loserMeanAce = mean(l_ace), 
    loserMeanDf = mean(l_df), 
    loserMeanSvpt = mean(l_svpt)
    ) 

atp_matches_2022_zscore <- atp_matches_2022_totals |> 
  mutate(
    winnerMeanAceZscore = as.numeric(scale(winnerMeanAce, center = TRUE, scale = TRUE)),
    winnerMeanDfZscore = as.numeric(scale(winnerMeanDf, center = TRUE, scale = TRUE)),
    winnerMeanSvptZscore = as.numeric(scale(winnerMeanSvpt, center = TRUE, scale = TRUE)),
    loserMeanAceZScore = as.numeric(scale(loserMeanAce, center = TRUE, scale = TRUE)) * -1,
    loserMeanDfZScore = as.numeric(scale(loserMeanDf, center = TRUE, scale = TRUE)) * -1,
    loserMeanSvptZScore = as.numeric(scale(loserMeanSvpt, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = winnerMeanAceZscore + winnerMeanDfZscore + winnerMeanSvptZscore + loserMeanAceZScore + loserMeanDfZScore + loserMeanSvptZScore
  ) |> filter(winner_name %in% male_winners_2022) |> arrange(desc(TotalZscore))

head(atp_matches_2022_zscore) 
```

```{r}
atp_matches_2023_totals <- atp_matches_2023_selected |> 
  group_by(winner_name) |> 
  na.omit() |>
  summarise(
    winnerMeanAce = mean(w_ace), 
    winnerMeanDf = mean(w_df), 
    winnerMeanSvpt = mean(w_svpt),
    loserMeanAce = mean(l_ace), 
    loserMeanDf = mean(l_df), 
    loserMeanSvpt = mean(l_svpt)
    ) 

atp_matches_2023_zscore <- atp_matches_2023_totals |> 
  mutate(
    winnerMeanAceZscore = as.numeric(scale(winnerMeanAce, center = TRUE, scale = TRUE)),
    winnerMeanDfZscore = as.numeric(scale(winnerMeanDf, center = TRUE, scale = TRUE)),
    winnerMeanSvptZscore = as.numeric(scale(winnerMeanSvpt, center = TRUE, scale = TRUE)),
    loserMeanAceZScore = as.numeric(scale(loserMeanAce, center = TRUE, scale = TRUE)) * -1,
    loserMeanDfZScore = as.numeric(scale(loserMeanDf, center = TRUE, scale = TRUE)) * -1,
    loserMeanSvptZScore = as.numeric(scale(loserMeanSvpt, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = winnerMeanAceZscore + winnerMeanDfZscore + winnerMeanSvptZscore + loserMeanAceZScore + loserMeanDfZScore + loserMeanSvptZScore
  ) |> filter(winner_name %in% male_winners_2023) |> arrange(desc(TotalZscore))

head(atp_matches_2023_zscore) 
```

```{r}
wta_matches_2022_totals <- wta_matches_2022_selected |> 
  group_by(winner_name) |> 
  na.omit() |>
  summarise(
    winnerMeanAce = mean(w_ace), 
    winnerMeanDf = mean(w_df), 
    winnerMeanSvpt = mean(w_svpt),
    loserMeanAce = mean(l_ace), 
    loserMeanDf = mean(l_df), 
    loserMeanSvpt = mean(l_svpt)
    ) 

wta_matches_2022_zscore <- wta_matches_2022_totals |> 
  mutate(
    winnerMeanAceZscore = as.numeric(scale(winnerMeanAce, center = TRUE, scale = TRUE)),
    winnerMeanDfZscore = as.numeric(scale(winnerMeanDf, center = TRUE, scale = TRUE)),
    winnerMeanSvptZscore = as.numeric(scale(winnerMeanSvpt, center = TRUE, scale = TRUE)),
    loserMeanAceZScore = as.numeric(scale(loserMeanAce, center = TRUE, scale = TRUE)) * -1,
    loserMeanDfZScore = as.numeric(scale(loserMeanDf, center = TRUE, scale = TRUE)) * -1,
    loserMeanSvptZScore = as.numeric(scale(loserMeanSvpt, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = winnerMeanAceZscore + winnerMeanDfZscore + winnerMeanSvptZscore + loserMeanAceZScore + loserMeanDfZScore + loserMeanSvptZScore
  ) |> filter(winner_name %in% female_winners_2022) |> arrange(desc(TotalZscore))

head(wta_matches_2022_zscore) 
```

```{r}
wta_matches_2023_totals <- wta_matches_2023_selected |> 
  group_by(winner_name) |> 
  na.omit() |>
  summarise(
    winnerMeanAce = mean(w_ace), 
    winnerMeanDf = mean(w_df), 
    winnerMeanSvpt = mean(w_svpt),
    loserMeanAce = mean(l_ace), 
    loserMeanDf = mean(l_df), 
    loserMeanSvpt = mean(l_svpt)
    ) 

wta_matches_2023_zscore <- wta_matches_2023_totals |> 
  mutate(
    winnerMeanAceZscore = as.numeric(scale(winnerMeanAce, center = TRUE, scale = TRUE)),
    winnerMeanDfZscore = as.numeric(scale(winnerMeanDf, center = TRUE, scale = TRUE)),
    winnerMeanSvptZscore = as.numeric(scale(winnerMeanSvpt, center = TRUE, scale = TRUE)),
    loserMeanAceZScore = as.numeric(scale(loserMeanAce, center = TRUE, scale = TRUE)) * -1,
    loserMeanDfZScore = as.numeric(scale(loserMeanDf, center = TRUE, scale = TRUE)) * -1,
    loserMeanSvptZScore = as.numeric(scale(loserMeanSvpt, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = winnerMeanAceZscore + winnerMeanDfZscore + winnerMeanSvptZscore + loserMeanAceZScore + loserMeanDfZScore + loserMeanSvptZScore
  ) |> filter(winner_name %in% female_winners_2023) |> arrange(desc(TotalZscore))

head(wta_matches_2023_zscore) 
```
Here I attempted to calculate the Z-scores for the top 20 grand slam players of the ATP and WTA from 2022 and 2023. What really surprised me is that none of the four names from the top of the prior bar and dumbbell charts were found in the top of the Z-score rankings. For example, we have some unknown players like Alexander Zverev and Felix Auger Aliassime who appear to be performing better than more recognized players like Rafael Nadal and Novak Djokovic. However, looking at the chart of overall matches, Zverev only won a total of eight grand slam matches in 2022, whereas the rest of them are comprised of matches won in non-GS tournaments. It's possible that some athletes put more of their focus on tournaments outside of the Grand Slam as to improve their overall record in tennis matches, whereas those who focus solely on the Grand Slams often train for the fame and glory. Or it's possible that this is because of the number of overall matches they have competed in throughout the year, as any higher number of matches won/loss could drastically sway their performance one way or another.